An Empirical Analysis of the Role of Energy in Economic Growth

### 2015-01-16 Articles/Article_EnerEcon_2014/EconGrowth4.Rnw
Finished boundary models.
Added text about CES boundary models.

The fit with lowest $sse$ of all above trials 
is deemed the winning model to be presented in graphs.

>>>>

****

In addition to the above trials, 
and in a manner similar to our approach to the Cobb-Douglas model,
we fit along all possible CES boundaries.
Boundaries are given by all combinations of the following:
$\delta_1$~=~0~or~1, 
$\sigma_1$~=~0~or~$\infty$,
$\delta$~=~0~or~1, and
$\sigma$~=~0~or~$\infty$.
Equation~\ref{eq:ces_generic} shows the generic three-factor CES equation,
wherein factors of production ($x_1$,~$x_2$,~$x_3$) 
are permutations of ($k$,~$l$,~$e$).
%
\begin{equation} \label{eq:ces_generic}
  y = \gamma A \left\{ \delta \left[ \delta_1 x_1^{-\rho_1} 
      + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{\rho/\rho_1}  
          + \left( 1 - \delta \right) x_3^{-\rho}  \right\} ^ {-1/\rho}
\end{equation}
%
Table~\ref{tab:CES_boundary_models} shows the set 
of degenerate equations found along parameter boundaries.
Each degenerate equation provides a boundary model.
Note that many boundary parameter combinations 
yield equivalent CES degenerate equations,%
  \footnote{
  For example, all possible boundary parameter combinations 
  that include $\delta = 0$ yield the same boundary model, $y = \gamma A x_3$.
  }
so there are far fewer than the maximum possible number of boundary models, 81.%
  \footnote{
  There are three possible states (unspecified, lower boundary, upper boundary)
  for four parameters ($\delta_1$, $\sigma_1$, $\delta$, $\sigma$),
  which gives $3^4 = 81$ possible boundary models.
  }
In the parameter columns of Table~\ref{tab:CES_boundary_models}, 
``---'' indicates that a parameter is unknowable on that boundary
(because it does not appear in the boundary model), and
an empty space indicates that the parameter 
can be estimated on that boundary using the degenerate equation.
The top portion of Table~\ref{tab:CES_boundary_models} shows 
two-parameter CES boundary models
(similar to the CES model without energy in Equation~\ref{eq:CES})
with $\delta = 1$, $\sigma$ unknowable, and $x_3$ absent.
If the fitted parameters in a boundary model violate constraints, 
the boundary model is rejected.

\begin{table} \caption{Boundary models for the CES equation.} 
\label{tab:CES_boundary_models} 
  \begin{center}
    \begin{tabular}{c c c c l} 
      \toprule
      $\delta_1$  & $\sigma_1$ & $\delta$   & $\sigma$ & Boundary model \\
      \midrule
      1           & ---        & 1          & ---      & $y = \gamma A x_1$ \\
      0           & ---        & 1          & ---      & $y = \gamma A x_2$ \\
      ---         & 0          & 1          & ---      & $y = \gamma A \min (x_1, \, x_2)$ \\
%
                  & $\infty$   & 1          & ---        
            & $y = \gamma A \left[ \delta_1 x_1 + (1 - \delta_1) x_2 \right]$ \\
%
                  &            & 1          & ---
            & $y = \gamma A \left[ \delta_1 x_1^{-\rho_1} 
                    + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{-1/\rho_1}$ \\
\midrule
      ---         & ---        & 0          & ---      & $y = \gamma A x_3$ \\
      1           & ---        & ---        & 0        & $y = \gamma A \min (x_1, \, x_3)$ \\
      0           & ---        & ---        & 0        & $y = \gamma A \min (x_2, \, x_3)$ \\
      ---         & 0          & ---        & 0        & $y = \gamma A \min (x_1, \, x_2, \, x_3)$ \\
%
      1           & ---        &            & $\infty$        
            & $y = \gamma A \left[ \delta x_1 + (1 - \delta) x_3 \right]$ \\
%
      0           & ---        &            & $\infty$        
            & $y = \gamma A \left[ \delta x_2 + (1 - \delta) x_3 \right]$ \\
%
                  & $\infty$   &            & $\infty$        
            & $y = \gamma A \left\{ \delta \left[ \delta_1 x_1 + \left( 1 - \delta_1 \right) x_2 \right] 
                    + \left( 1 - \delta \right) x_3  \right\}$ \\
%
      ---         & 0          &            & $\infty$        
            & $y = \gamma A \left[ \delta \min (x_1, \, x_2) + \left( 1 - \delta \right) x_3  \right]$ \\
%
                  & $\infty$   & ---        & 0
            & $y = \gamma A \min \left[ \delta_1 x_1 + \left(1 - \delta_1 \right) x_2, \, x_3  \right]$ \\
%
      1           & ---        &         & 
            & $y = \gamma A \left[ \delta x_1^{-\rho} 
                    + \left(1 - \delta \right) x_3^{-\rho} \right] ^ {-1/\rho}$ \\ 
%
      0           & ---        &            & 
            & $y = \gamma A \left[ \delta x_2^{-\rho} + \left(1 - \delta \right) x_3^{-\rho} \right]^{-1/\rho}$ \\
%
                  &            & ---     & 0
            & $y = \gamma A \min \left\{ \left[ \delta_1 x_1^{-\rho_1} 
                    + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{-1/\rho_1}, \, x_3 \right\}$ \\ 
%
      ---         & 0          &            & 
            & $y = \gamma A \left\{ \delta \left[ \min \left( x_1, \, x_2 \right) \right]^{-\rho}  
                    + \left( 1 - \delta \right) x_3^{-\rho}  \right\} ^ {-1/\rho}$ \\ 
%
                  &            &            & $\infty$
            & $y = \gamma A \left\{ \delta \left[ \delta_1 x_1^{-\rho_1} 
                    + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{-1/\rho_1}  
                    + \left( 1 - \delta \right) x_3  \right\}$ \\ 
%
                  & $\infty$   &            &
            & $y = \gamma A \left\{ \delta \left[ \delta_1 x_1 
                    + \left( 1 - \delta_1 \right) x_2 \right]^{-\rho}  
                    + \left( 1 - \delta \right) x_3^{-\rho}  \right\} ^ {-1/\rho}$ \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

****

The fit with lowest $sse$ of all above trials 
is deemed the winning model and presented in graphs.
